\chapter[Include Files]{Include Files} \label{chap:inc_files}

The files presented here are located in {\em Config/} directory.
They are required for the preprocessing and compilation of the
code. 
%Any modification of these files leads to the execution of the
%commands {\tt mkmf} for the creation of the Makefile file and
%{\tt make} in order to obtain a new executable.

\vskip 0.5cm
\noindent
{\em gem\_config.h} ,{\em compiler.mk} and {\em libraries.mk}
\newline
Sets configuration parameters for gem Makefiles.
The following information must be provided:
\begin{itemize}
\item Location of the netCDF include files (variable {\em INCLUDES\_NETCDF}) 
      and library ({\em NETCDFLibs})
\item Location of the ESMF include files (variable {\em INCLUDES\_Esmf}) 
      and library ({\em EsmfLibs})
\item Location of the MPI include files (variable {\em INCLUDES\_MSG})
      and library ({\em MSGLIBDIR})
\item Commands to call the C (variable {\em CC}) and Fortran ({\em FC})
      compilers.
\item Compilations options.
\end{itemize}


\vskip 0.5cm
\noindent
{\bf gem\_options.h}
\newline
To select the package, the chemical solver, and the desire
compilation mode (debugging, optimization, profiling).


\vskip 0.5cm
\noindent
{\bf gem\_sys\_options.h}
\newline
To select the architecture and the message passing options.
The user must set the variables:
\begin{itemize}
\item {\em ARCH\_OPTION}: to determine the platform used.
\item {\em MSG\_OPTION}: for the message passing option.
      Choose {\em MSG\_NONE} if you want the single processor
      version of the code, or {\em MSG\_MPI} if you prefer
      the multiple processor one with MPI as message passing.
\item {\em MPI\_2\_OPTION}: to determine if you want to include
      or not MPI-2 calls. Consider {\em NO\_MPI\_2} if the platform
      you will be running the code on does not support MPI-2.
      Otherwise choose {\em WITH\_MPI\_2}.
\end{itemize}

\vskip 0.5cm
\noindent
{\bf gem\_msg\_numbers.h}
\newline
Contains the message numbers used to identify specific
messages in the message passing calls.
Should not be edited.

\vskip 0.5cm
\noindent
{\bf gem\_rules.h} and {\em rules.mk}
\newline
Contains suffix dependencies and compilation rules for all
Makefile/Makefile.cpp files in directories that contain source files.
Should not be edited.

\vskip 0.5cm
\noindent
With time, we will make necessary changes to remove all the ".h" files and
only keep the ".mk" ones.
